Refinements for open automata

### 2021-11-04 2021-Refinement/paper/refinementarticle.tex
vars in ots

\usepackage{tikz}

@@

- a first relation that focuses on the automaton part and a second that deals partially with open behaviour

we will see that having at the same time composition and transitivity also raises challenges; we introduce a relation that addresses this challenge.

@@

The notion of refinement aims of capturing the relation between  a specification and an implementation of the same component. It is usually defined as trace inclusion or simulation \cite{Milner:1980, Kouchnarenko:2007}; this ensures that all behaviours of the implementation must be also behaviours of the specification. This definition, which is based on systems whose behaviour is fully defined, is not well-suited for open systems,  as it requires to reason also about unspecified behaviours.

@@

An \emph{open transition} is a tuple \nmm{\OTg} with \(s, s' \in S\) the source and target states, \(\alpha \in \actions\) the produced action, \(J' \subseteq J\) the holes involved in the transition, \(\beta_j \in \actions\) the actions of the holes, \(g \in \formulas\) the guard and \(\psi: V \to \terms\) the variable assignments.
\end{definition}

%\TODO{useful or not?}
\begin{definition}[Guard, Out-transition, Transition variables]
Let \(V\) be the variable names of the considered automaton, \(T\) its transitions and \(r\) one of its states.
\(\fOT[T]{r} \in T\) are called the out-transitions of the state \(r\).
\(\fIT[T]{r} \in T\) are called the in-transitions of the state \(r\).
When the transition set is clear from the context, it will be omitted.
The local variables of a transition \(\fvars{t}\) are all variables appearing in transition \(t\) except the global variables of the automaton.
\begin{align*}
	\fOT[T]{r} & = \mset{\OTg \in T}{s = r} &
	\fIT[T]{r} & = \mset{\OTg \in T}{s' = r} \\
\end{align*}
\vspace{-1cm}
\begin{gather*}
	\fguard{\OTg} = g \\
	\fvars{\OTg} = \mpar{\fvars{\alpha} \cup \fvars{g} \cup \bigcup_{j \in J'} \fvars{\beta_j} \cup \bigcup_{v \in V} \fvars{\psi\mpar{v}}} \setminus V
\end{gather*}
\end{definition}

@@

The composition \(A_0\mdbrk{j_i \mapsto A_i \middle| 1 \leq i \leq n}\) is non-locking if   \(A \) has a reachability predicate such that, for each reachable configuration, if there is a possible transition in \(A_0\) then there is a possible transition in \(A\):
\[ \forall s = \mpar{s_0, s_1,..,s_n} \in S, V \uplus \biguplus_\subbox{t_0 \in \fOT{s_0}} \fvars{t_0} \vdash \left(\reach{A}\mpar{s} \wedge \bigvee_\subbox{t_0 \in \fOT{s_0}} \fguard{t_0} \implies \bigvee_\subbox{t \in \fOT{s}} \fguard{t} \right)\]
\end{definition}
Again we use guards to characterise that the transition occurs. It is not sufficient to ensure the existence of equivalent transitions in general, but it will be sufficient in the context of refinement.

>>>>

\usepackage{tikz,mathpartir}

@@

- a first relation that focuses on the automaton part 
and a second that deals with open behaviour

we will see that having at the same time composition and transitivity also raises challenges; we introduce a new form of refinement relation that addresses this challenge.

@@

The notion of refinement aims captures the relation between  a specification and an implementation of the same component. It is usually defined as trace inclusion or simulation \cite{Milner:1980, Kouchnarenko:2007}; this ensures that all behaviours of the implementation must be also behaviours of the specification. This definition, which is based on systems whose behaviour is fully defined, is not well-suited for open systems,  as it requires to reason also about unspecified behaviours.

@@

An \emph{open transition} is a tuple \nmm{\OTg} with \(s, s' \in S\) the source and target states, \(\alpha \in \actions\) the produced action, \(J' \subseteq J\) the holes involved in the transition, \(\beta_j \in \actions\) the actions of the holes, \(g \in \formulas\) the guard and \(\psi: V \to \terms\) the variable assignments.
To be well-formed, an open transition should use only variables of the automaton and variables appearing in the involved actions, formally: 
\begin{align*}
\fvars(g)&\subseteq \fvars(\alpha)\cup \bigcup_{j \in J'} \fvars{\beta_j} \cup V \\ \forall v\in V.\, \fvars(\psi(v))&\subseteq \fvars(\alpha)\cup \bigcup_{j \in J'} \fvars{\beta_j} \cup V
\end{align*}
\end{definition}

%\TODO{useful or not?}
\begin{definition}[Guard, Out-transition, Transition variables]
Let \(V\) be the variable names of the considered automaton, \(T\) its transitions and \(r\) one of its states.
\(\fOT[T]{r} \in T\) are called the out-transitions of the state \(r\).
\(\fIT[T]{r} \in T\) are called the in-transitions of the state \(r\).
When the transition set is clear from the context, it will be omitted.
The local variables of a transition \(\fvars{t}\) are all variables appearing in transition \(t\) except the global variables of the automaton.
\begin{align*}
	\fOT[T]{r} & = \mset{\OTg \in T}{s = r} &
	\fIT[T]{r} & = \mset{\OTg \in T}{s' = r} \\
\end{align*}
\vspace{-1cm}
\begin{mathpar}
	\fguard{\OTg} \!=\! g \qquad
	\fvars{\OTg} \!=\! \mpar{\fvars{\alpha}  \cup \bigcup_{j \in J'} \!\fvars{\beta_j} } \setminus V
\end{mathpar}
\end{definition}

@@

The composition \(A_0\mdbrk{j_i \mapsto A_i \middle| 1 \leq i \leq n}\) is non-locking if   \(A \) has a reachability predicate such that, for each reachable configuration, if there is a possible transition in \(A_0\) then there is a possible transition in \(A\):
\[ \forall s = \mpar{s_0, s_1,..,s_n} \in S, V \uplus \biguplus_\subbox{t_0 \in \fOT{s_0}} \fvars{t_0} \vdash \left(\reach{A}\mpar{s} \wedge \bigvee_\subbox{t_0 \in \fOT{s_0}} \fguard{t_0} \implies \bigvee_\subbox{t \in \fOT{s}} \fguard{t} \right)\]
\end{definition}
Again we use guards to ensure that the transition can occur. It is not sufficient to ensure the existence of equivalent transitions in general, but it will be sufficient in the context of refinement.
